---
title: "ACT Data Prep 2025"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
---

Author: Mengna Zhang


*Last updated on:* `r Sys.Date()`

# Set Path

```{r set-path}
# Clear the workspace to ensure a clean environment
rm(list = ls())

# Clear plots and console history
if (!is.null(dev.list())) dev.off()  # close any open plotting devices

## directory: can extend to the main CNT folder
directory <- "/Users/mengnazhang/Desktop/ADSP_DataPrep_local/"

## set PRADI path (the folder where PRADI raw files located)
act_directory <- paste0(directory,"ACT/Phenotype/2025/Raw/")

## output path
out_directory <- paste0(directory,"ACT/Phenotype/2025/Cleaned/")

## script path
script_directory <- paste0(directory,"ACT/Phenotype/2025/Scripts/")

```

<br>

# Load Helper Scripts
```{r load-helper-scripts}
source(paste0(script_directory,"helperScripts_ACT.R"))
```

<br>

# Load Packages
```{r laod-packages,message=FALSE,warning=FALSE}
require(readxl)
require(DT)
```

<br>

# Explore Data Dictionary

**Purpose**: Import the data dictionary and investigate how variables and codes are defined. This exploration will inform the creation of validation functions to flag invalid or inconsistent values within the datasets.

## Load DD
```{r exploreDD-1}
DD_path <- paste0(act_directory,"Frz2024_DataDictionary_2025_06_06.xlsx")

DD_sheet_names <- excel_sheets(DD_path)

print(DD_sheet_names) ## 106 sheets in total

## remove the "Table of Contents" from sheet_names, we dont need to read it
DD_sheet_names <- setdiff(DD_sheet_names,"Table of Contents")

## read each sheet, skipping the first 5 rows, since they are meta info, not relevant
all_DD_sheets <- lapply(DD_sheet_names, function(sheet) {
  read_excel(DD_path, sheet = sheet, skip = 5)
})


## load sheet names of datasets and remove the DD sheets not present in the datasets
df_path <- paste0(act_directory,"ACT_353_Hohman_2025_09_11.xlsx")

df_sheet_names <- excel_sheets(df_path)

print(df_sheet_names) ## 32 sheets in total

## replace " - " with "_" in DD_sheet_names
DD_sheet_names <- gsub(" - ", "_", DD_sheet_names)

## check what other names should be modified in the DD_sheet_names:
setdiff(df_sheet_names,DD_sheet_names)
## [1] "Form78_InjuryLOC" "genetics_apoe"    "autopsy_status"   "np_standard"      "nic_cdes"         "nic_am_vols"     
## [7] "nic_pm_vols"   

DD_sheet_names[DD_sheet_names == "Form78_InjuryLoc"] <- "Form78_InjuryLOC"
DD_sheet_names[DD_sheet_names == "Genetics_APOE"] <- "genetics_apoe"
DD_sheet_names[DD_sheet_names == "Autopsy_Status"] <- "autopsy_status"
DD_sheet_names[DD_sheet_names == "Neuropathology_Standard"] <- "np_standard"
DD_sheet_names[DD_sheet_names == "Neuroimaging_CDEs"] <- "nic_cdes"
DD_sheet_names[DD_sheet_names == "Neuroimaging_AM_Vols"] <- "nic_am_vols"
DD_sheet_names[DD_sheet_names == "Neuroimaging_PM_Vols"] <- "nic_pm_vols"

setdiff(df_sheet_names,DD_sheet_names) ## character(0)

# name the DD sheet names list
names(all_DD_sheets) <- DD_sheet_names

## subset DD sheet names to be consistent with datasets
DD_sheet_names <- intersect(DD_sheet_names,df_sheet_names)

## remove irrelevant elements from the all_DD_sheets list
all_DD_sheets <- all_DD_sheets[DD_sheet_names] ## length = 32

## extract the Coding variables acorss all the DDs and check unique values
all_coding_df <- data.frame(
  Coding = unlist(lapply(all_DD_sheets, function(df) df[["Coding"]])),
  Type   = unlist(lapply(all_DD_sheets, function(df) df[["Type"]])),
  stringsAsFactors = FALSE
) 

# remove NAs and duplicates (optional)
all_coding_df <- unique(na.omit(all_coding_df)) ## 157 obs
all_coding_df_chr <- all_coding_df[all_coding_df$Type == "Char",] ## 7 obs
all_coding_df_num <- all_coding_df[all_coding_df$Type == "Num",] ## 150 obs
```


<br>

## Functions: Check Valid Values

After observing the patterns of Coding across all the datasets, there are some fixed patterns, where xx represents random texts:
* "1 = xx\r\n2 = xx\r\n3 = xx\r\n. = xx"
  * Valid Values: 1,2,3
  * Strategy: remove "\r\n", and retrive unique numbers of the left side of "=" sign

* "Integer from 0 through 50\r\n88 = xx\r\n99 = xx\r\n. = xx"
  * Valid Values: integer, 0-50,88,99
* "Non-negative integers\r\n88 = xx\r\n99 = Txxr\r\n. = xx" 
  * Valid Values: integer, >0, 88, 99
* "Non-negative integers"
* "Positive integers, round to integer\r\n"
  * strategy: if detected from .. through .. then extract the range. If detected "Integer" before that, make sure the data type is integer not numeric, if "=" presents, then also retrive the number on the left side of "=" sign. Finally, merge all the values range and unique values together into a vector. If found non-negative keywords, then make sure the value > 0 + data type = integer

* "Positive real numbers, round to 2 decimals\r\n"
* "Positive real numbers (to 1 decimal)\r\n. = Missing"
* "Positive real numbers, round to 2 decimals\r\n. = Missing"
* "Positive real numbers, millimeters (mm)\r\n. = Missing"
* "Real numbers\r\n. = Missing"
  * strategy: if detected "real numbers", then make sure the data type is numeric, if detected positive then make sure the value > 0

* "Valid years starting in 1994"
* "Valid ages 65+"
* "Valid years\r\n. 
  * strategy: if detected "valid years|ages", then make sure the value is positive, if additional number is provided, then make sure the values greated or equal that value


* "Number of years\r\n. = Missing" ==> valid values: > 0
  * strategy: if detected "number of years" than make sure the data value > 0

<br>

### Core Function
```{r exploreDD-2}
extract_valid_values <- function(text) {
  if (is.na(text) || trimws(text) == "") {
    return(list(type = NA, valid_values = NA))
  }
  
  txt <- tolower(gsub("\r\n", " ", text))
  
  result <- list(type = NA, valid_values = NA)
  values <- numeric()
  range_vals <- NULL
  range_str <- NULL
  min_val <- NULL
  
  # ---- Extract explicit equality numbers ----
  eq_nums <- regmatches(txt, gregexpr("\\b\\d+\\s*=", txt))
  if (length(eq_nums[[1]]) > 0) {
    nums <- as.numeric(gsub("\\s*=.*", "", eq_nums[[1]]))
    values <- unique(nums)
  }
  
  # ---- Detect range patterns ----
  # Handles:
  #   "from 0 through 6", "from 0 to 6", "from 0-6"
  #   "between 0 and 6", "integers between 0 and 6"
  #   "0-6"
  range_pattern <- "(from|between)?\\s*\\d+(\\s*(through|to|and|-|–)\\s*\\d+)"
  if (grepl(range_pattern, txt)) {
    nums <- as.numeric(unlist(regmatches(txt, gregexpr("\\d+(?:\\.\\d+)?", txt))))
    if (length(nums) >= 2) {
      range_vals <- nums[1]:nums[2]
      range_str <- paste0(nums[1], "-", nums[2])
    }
  }
  
  
  # ---- Detect keywords for type ----
  if (grepl("integer", txt)) result$type <- "integer"
  else if (grepl("real", txt)) result$type <- "numeric"
  
  # ---- Detect positive / non-negative ----
  if (grepl("non[- ]negative", txt)) min_val <- 0
  if (grepl("positive", txt)) min_val <- 0
  
  # ---- Detect "valid years/dates starting in X" ----
  if (grepl("valid (years|dates) starting in", txt)) {
    yr <- as.numeric(gsub(".*valid (years|dates) starting in\\s*(\\d+).*", "\\2", txt))
    if (!is.na(yr)) min_val <- yr
  }
  
  # ---- Detect "valid ages X+" ----
  if (grepl("valid ages", txt)) {
    age <- as.numeric(gsub(".*valid ages\\s*(\\d+)\\+.*", "\\1", txt))
    if (!is.na(age)) min_val <- age
  }
  
  # ---- Detect "number of years" ----
  if (grepl("number of years", txt)) min_val <- 0
  
  # ---- Build valid_values string ----
  valid_str <- NULL
  
  # Case 1: min value detected (>=)
  if (!is.null(min_val)) {
    valid_str <- paste0(">=", min_val)
  }
  
  # Case 2: range detected
  if (!is.null(range_str)) {
    if (length(values) > 0) {
      # only include values outside the range
      outside_vals <- setdiff(values, range_vals)
      if (length(outside_vals) > 0) {
        valid_str <- paste0(range_str, ",", paste(sort(outside_vals), collapse = ","))
      } else {
        valid_str <- range_str
      }
    } else {
      valid_str <- range_str
    }
  } 
  
  # Case 3: discrete values only
  if (is.null(range_str) && length(values) > 0 && is.null(valid_str)) {
    valid_str <- paste(sort(unique(values)), collapse = ",")
  }
  
  # ---- Infer type if missing ----
  if (is.na(result$type)) {
    if (!is.null(valid_str) && grepl("^[0-9,\\-]+$", valid_str)) {
      result$type <- "integer"
    } else {
      result$type <- "numeric"
    }
  }
  
  result$valid_values <- valid_str
  return(result)
}

```

<br>

### Core Function Validation
```{r exploreDD-3}
## apply the function to the all_coding_df_num dataset to validate the function
generate_valid_values <- function(df) {
  # Apply the extraction function to each Coding cell
  parsed <- lapply(df$Coding, extract_valid_values)
  
  # Extract the two elements (type, valid_values) into separate columns
  df$type <- sapply(parsed, function(x) {
    ifelse(is.null(x$type) || is.na(x$type), NA, x$type)
  })
  
  df$valid_values <- sapply(parsed, function(x) {
    ifelse(is.null(x$valid_values) || is.na(x$valid_values), NA, x$valid_values)
  })
  
  return(df)
}

tmp <- generate_valid_values(all_coding_df_num)

DT::datatable(tmp)
```

