---
title: "ACT Data Prep 2025"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
---

Author: Mengna Zhang


*Last updated on:* `r Sys.Date()`

# Set Path

```{r set-path}
# Clear the workspace to ensure a clean environment
rm(list = ls())

# Clear plots and console history
if (!is.null(dev.list())) dev.off()  # close any open plotting devices

## directory: can extend to the main CNT folder
directory <- "/Users/mengnazhang/Desktop/ADSP_DataPrep_local/"

## set PRADI path (the folder where PRADI raw files located)
act_directory <- paste0(directory,"ACT/Phenotype/2025/Raw/")

## output path
out_directory <- paste0(directory,"ACT/Phenotype/2025/Cleaned/")

## script path
script_directory <- paste0(directory,"ACT/Phenotype/2025/Scripts/")

```

<br>

# Load Helper Scripts
```{r load-helper-scripts}
source(paste0(script_directory,"helperScripts_ACT.R"))
```

<br>

# Load Packages
```{r laod-packages,message=FALSE,warning=FALSE}
require(readxl)
require(DT)
require(dplyr)
```

<br>

# Explore Data Dictionary

**Purpose**: Import the data dictionary and investigate how variables and codes are defined. This exploration will inform the creation of validation functions to flag invalid or inconsistent values within the datasets.

## Load DD
```{r exploreDD-1}
DD_path <- paste0(act_directory,"Frz2024_DataDictionary_2025_06_06.xlsx")

DD_sheet_names <- excel_sheets(DD_path)

print(DD_sheet_names) ## 106 sheets in total

## remove the "Table of Contents" from sheet_names, we dont need to read it
DD_sheet_names <- setdiff(DD_sheet_names,"Table of Contents")

## read each sheet, skipping the first 5 rows, since they are meta info, not relevant
all_DD_sheets <- lapply(DD_sheet_names, function(sheet) {
  read_excel(DD_path, sheet = sheet, skip = 5)
})


## load sheet names of datasets and remove the DD sheets not present in the datasets
df_path <- paste0(act_directory,"ACT_353_Hohman_2025_09_11.xlsx")

df_sheet_names <- excel_sheets(df_path)

print(df_sheet_names) ## 32 sheets in total

## replace " - " with "_" in DD_sheet_names
DD_sheet_names <- gsub(" - ", "_", DD_sheet_names)

## check what other names should be modified in the DD_sheet_names:
setdiff(df_sheet_names,DD_sheet_names)
## [1] "Form78_InjuryLOC" "genetics_apoe"    "autopsy_status"   "np_standard"      "nic_cdes"         "nic_am_vols"     
## [7] "nic_pm_vols"   

DD_sheet_names[DD_sheet_names == "Form78_InjuryLoc"] <- "Form78_InjuryLOC"
DD_sheet_names[DD_sheet_names == "Genetics_APOE"] <- "genetics_apoe"
DD_sheet_names[DD_sheet_names == "Autopsy_Status"] <- "autopsy_status"
DD_sheet_names[DD_sheet_names == "Neuropathology_Standard"] <- "np_standard"
DD_sheet_names[DD_sheet_names == "Neuroimaging_CDEs"] <- "nic_cdes"
DD_sheet_names[DD_sheet_names == "Neuroimaging_AM_Vols"] <- "nic_am_vols"
DD_sheet_names[DD_sheet_names == "Neuroimaging_PM_Vols"] <- "nic_pm_vols"

setdiff(df_sheet_names,DD_sheet_names) ## character(0)

# name the DD sheet names list
names(all_DD_sheets) <- DD_sheet_names

## subset DD sheet names to be consistent with datasets
DD_sheet_names <- intersect(DD_sheet_names,df_sheet_names)

## remove irrelevant elements from the all_DD_sheets list
all_DD_sheets <- all_DD_sheets[DD_sheet_names] ## length = 32

## extract the Coding variables acorss all the DDs and check unique values
all_coding_df <- do.call(rbind, lapply(names(all_DD_sheets), function(sheet_name) {
  df <- all_DD_sheets[[sheet_name]]
  
  # Ensure we only extract existing columns
  var <- if ("Variable" %in% names(df)) df$Variable else NA
  coding <- if ("Coding" %in% names(df)) df$Coding else NA
  type <- if ("Type" %in% names(df)) df$Type else NA
  
  # Repeat Df_name for each row of this tibble
  data.frame(
    Df_name = rep(sheet_name, length(var)),
    Variable = var,
    Coding = coding,
    Type = type,
    stringsAsFactors = FALSE
  )
}))

dim(all_coding_df) ## [1] 1374    4

## split by type
unique(all_coding_df$Type)

## explore NA and fill the missingness
print(all_coding_df[is.na(all_coding_df$Type),])
## Type = Num for those 3 NA
all_coding_df$Type[is.na(all_coding_df$Type)] <- "Num"

all_coding_df_chr <- all_coding_df[all_coding_df$Type == "Char",] ## 18 obs
length(unique(all_coding_df_chr$Coding)) ## 7 unique patterns in the DD

all_coding_df_num <- all_coding_df[all_coding_df$Type == "Num",] ## 1359 obs
length(unique(all_coding_df_num$Coding)) ## 150 unique patterns in the DD

```




<br>

## Functions: Extract Valid Values for Numeric Variables

Patterns observed from DD

* "1 = xx\r\n2 = xx\r\n3 = xx\r\n. = xx"

* "Integer from 0 through 50\r\n88 = xx\r\n99 = xx\r\n. = xx"

* "Non-negative integers\r\n88 = xx\r\n99 = Txxr\r\n. = xx" 

* "Non-negative integers"

* "Positive integers, round to integer\r\n"
 
* "Positive real numbers, round to 2 decimals\r\n"

* "Positive real numbers (to 1 decimal)\r\n. = Missing"

* "Positive real numbers, round to 2 decimals\r\n. = Missing"

* "Positive real numbers, millimeters (mm)\r\n. = Missing"

* "Real numbers\r\n. = Missing"

* "Valid years starting in 1994"

* "Valid ages 65+"

* "Valid years\r\n.

* "Number of years\r\n. = Missing" ==> valid values: > 0

<br>

### Core Function
```{r exploreDD-2}
extract_valid_values <- function(text) {
  if (is.na(text) || trimws(text) == "") {
    return(list(type = NA, valid_values = NA))
  }
  
  txt <- tolower(gsub("\r\n", " ", text))
  
  result <- list(type = NA, valid_values = NA)
  values <- numeric()
  range_vals <- NULL
  range_str <- NULL
  min_val <- NULL
  
  # ---- Extract explicit equality numbers ----
  # Capture things like "1 = ", "10=", etc.
  eq_nums <- regmatches(txt, gregexpr("\\b\\d+\\s*=", txt))
  if (length(eq_nums[[1]]) > 0) {
    nums <- as.numeric(gsub("\\s*=.*", "", eq_nums[[1]]))
    values <- unique(nums)
  }
  
  # ---- Detect range patterns (outside code definitions) ----
  # Remove code definitions first so we don't confuse description ranges
  txt_no_eq <- gsub("\\b\\d+\\s*=.*?(?=\\b\\d+\\s*=|$)", "", txt, perl = TRUE)
  
  range_pattern <- "(from|between)?\\s*\\d+(\\s*(through|to|and|-|–)\\s*\\d+)"
  if (grepl(range_pattern, txt_no_eq)) {
    nums <- as.numeric(unlist(regmatches(txt_no_eq, gregexpr("\\d+(?:\\.\\d+)?", txt_no_eq))))
    if (length(nums) >= 2) {
      range_vals <- nums[1]:nums[2]
      range_str <- paste0(nums[1], "-", nums[2])
    }
  }
  
  # ---- Detect keywords for type ----
  if (grepl("integer", txt)) result$type <- "integer"
  else if (grepl("real", txt)) result$type <- "numeric"
  
  # ---- Detect positive / non-negative ----
  if (grepl("non[- ]negative", txt)) min_val <- 0
  if (grepl("positive", txt)) min_val <- 0
  
  # ---- Detect "valid years/dates starting in X" ----
  if (grepl("valid (years|dates) starting in", txt)) {
    yr <- as.numeric(gsub(".*valid (years|dates) starting in\\s*(\\d+).*", "\\2", txt))
    if (!is.na(yr)) min_val <- yr
  }
  
  # ---- Detect "valid ages X+" ----
  if (grepl("valid ages", txt)) {
    age <- as.numeric(gsub(".*valid ages\\s*(\\d+)\\+.*", "\\1", txt))
    if (!is.na(age)) min_val <- age
  }
  
  # ---- Detect "number of years" ----
  if (grepl("number of years", txt)) min_val <- 0
  
  # ---- Build valid_values string ----
  valid_str <- NULL
  
  # Case 1: min value detected (>=)
  if (!is.null(min_val)) {
    valid_str <- paste0(">=", min_val)
  }
  
  # Case 2: range detected (outside of code defs)
  if (!is.null(range_str)) {
    if (length(values) > 0) {
      outside_vals <- setdiff(values, range_vals)
      if (length(outside_vals) > 0) {
        valid_str <- paste0(range_str, ",", paste(sort(outside_vals), collapse = ","))
      } else {
        valid_str <- range_str
      }
    } else {
      valid_str <- range_str
    }
  } 
  
  # Case 3: discrete values only
  if (is.null(range_str) && length(values) > 0 && is.null(valid_str)) {
    valid_str <- paste(sort(unique(values)), collapse = ",")
  }
  
  # ---- Infer type if missing ----
  if (is.na(result$type)) {
    if (!is.null(valid_str) && grepl("^[0-9,\\-]+$", valid_str)) {
      result$type <- "integer"
    } else {
      result$type <- "numeric"
    }
  }
  
  result$valid_values <- valid_str
  return(result)
}

```

<br>

### Core Function Validation
```{r exploreDD-3}
## apply the function to the all_coding_df_num dataset to validate the function
generate_valid_values <- function(df) {
  # Apply the extraction function to each Coding cell
  parsed <- lapply(df$Coding, extract_valid_values)
  
  # Extract the two elements (type, valid_values) into separate columns
  df$type <- sapply(parsed, function(x) {
    ifelse(is.null(x$type) || is.na(x$type), NA, x$type)
  })
  
  df$valid_values <- sapply(parsed, function(x) {
    ifelse(is.null(x$valid_values) || is.na(x$valid_values), NA, x$valid_values)
  })
  
  return(df)
}

numValidVals <- generate_valid_values(all_coding_df_num)

DT::datatable(unique(numValidVals[,c("Coding","Type","type","valid_values")]))
```

<br>

## Functions: Extract Valid Values for Char Variables

Since there are only 7 character variables based on the DD. I will manually extract the valid values as above.

```{r exploreDD-4}
chrValidVals <- all_coding_df_chr %>%
  mutate(type = "character",
         valid_values = case_when(
    Coding == "C = Clinical\r\nR = Research\r\nE = ExVivo" ~ "C,R,E",
    Coding == "Text string of either 'Complete' or 'Unusable'" ~ "Complete,Unusable",
    Coding == "Text string of either 'Complete', 'Pending', or 'Unusable'" ~ "Complete,Pending,Unusable",
    Coding == "L = Left Hemisphere\r\nR = Right Hemisphere\r\nW = Whole Brain" ~ "L,R,W"
  ))
```


```{r exploreDD-5}
## combine
validVals <- rbind(numValidVals,chrValidVals)

```


<br>


# Load Data

```{r loadDD-1}
for (sheet in df_sheet_names) {
  assign(sheet, read_excel(df_path, sheet = sheet), envir = .GlobalEnv)
}
```



<br>

# Variable Check Per File

## Functions: Check Valid Values

```{r varCheck-1}
validate_dataset <- function(dataset_name, IDcol = "ID") {

  df <- get(dataset_name)
  
  # ---- Variable coverage check ----
  df_vars <- setdiff(names(df), IDcol)
  dd_vars <- validVals$Variable[validVals$Df_name == dataset_name]
  missed <- setdiff(df_vars, dd_vars)
  common_vars <- intersect(df_vars, dd_vars)
  
  # cat("\n df_vars are: ",df_vars,"\n")
  # cat("\n dd_vars are: ",dd_vars,"\n")
  # cat("\n commonDf are: ",commonDf,"\n")
  
  cat("\n---------------- Variable Coverage Check ----------------\n")
  if (length(missed) == 0) {
    cat("All variables in", dataset_name, "were found in the data dictionary.\n")
  } else {
    cat("Variables missing from data dictionary for", dataset_name, ":\n  ",
        paste(missed, collapse = ", "), "\n")
  }
  cat("----------------------------------------------------------\n\n")

  # ---- 3. Extract relevant rules from validVals ----
  rules <- validVals[validVals$Df_name %in% dataset_name,]

  results <- list()
  
  # print(rules)
  
  # ================================================================
  # 4. Loop through each variable defined in both dataset and dictionary
  # ================================================================
  for (varname in common_vars) {

    vals <- df[[varname]]
    obs_type <- class(df[[varname]])

    # ---- Extract expected information from dictionary ----
    exp_type <- rules[rules$Variable == varname,"type"]
    exp_values <- rules[rules$Variable == varname,"valid_values"]
    coding_pattern <- rules[rules$Variable == varname,"Coding"]

    # Initialize placeholders
    invalid_vals <- NA
    expected_vals <- NA
    observed_vals_str <- NA

    # ================================================================
    # 5. Convert variable to expected type (based on exp_type)
    # ================================================================
    if (!is.na(exp_type)) {
      if (exp_type == "character") {
        suppressWarnings( df[[varname]] <- as.character(df[[varname]]))
      } else if (exp_type == "numeric" && !is.numeric(vals)) {
        suppressWarnings( df[[varname]] <- as.numeric(df[[varname]]))
      } else if (exp_type == "integer" && !is.character(vals)) {
        suppressWarnings( df[[varname]] <- as.integer(df[[varname]]))
      }
    }


    # ================================================================
    # 6. Validation logic
    # ================================================================
    has_valid_values <- !is.na(exp_values) && trimws(exp_values) != ""

    if (has_valid_values) {
      uv <- unique(na.omit(vals))

      # ---- Numeric/integer checks ----
      if (tolower(exp_type) %in% c("numeric", "integer")) {
        if (grepl("^>=", exp_values)) {
          minval <- as.numeric(gsub("[^0-9.-]", "", exp_values))
          invalid_vals <- uv[uv < minval]
          expected_vals <- paste0(">= ", minval)

        } else if (grepl("^\\d+-\\d+$", exp_values)) {
          nums <- as.numeric(unlist(strsplit(exp_values, "-")))
          allowed <- nums[1]:nums[2]
          invalid_vals <- setdiff(uv, allowed)
          expected_vals <- paste0(nums[1], "-", nums[2])

        } else if (grepl("^[0-9,\\-]+$", exp_values)) {
          parts <- unlist(strsplit(exp_values, ","))
          allowed <- c()
          for (p in parts) {
            if (grepl("-", p)) {
              rng <- as.numeric(unlist(strsplit(p, "-")))
              allowed <- c(allowed, seq(rng[1], rng[2]))
            } else {
              allowed <- c(allowed, as.numeric(p))
            }
          }
          allowed <- unique(allowed)
          invalid_vals <- setdiff(uv, allowed)
          expected_vals <- paste(sort(allowed), collapse = ",")
        } else {
          expected_vals <- exp_values
        }

        # ---- Character checks ----
      } else if (tolower(exp_type) == "character") {
        allowed <- trimws(unlist(strsplit(exp_values, ",")))
        invalid_vals <- setdiff(unique(na.omit(vals)), allowed)
        expected_vals <- paste(allowed, collapse = ",")
      }

      # Limit observed values display to first 8
      observed_vals_str <- if (length(uv) > 8)
        paste(c(head(sort(uv), 8), "..."), collapse = ",")
      else paste(sort(uv), collapse = ",")

    } else {
      # ---- Missing valid_values → manual check needed ----
      expected_vals <- coding_pattern
      invalid_vals <- "Need manual check"

      if (is.numeric(vals)) {
        rng <- range(vals, na.rm = TRUE)
        observed_vals_str <- paste0(rng[1], "-", rng[2])
      } else if (is.character(vals)) {
        uv <- unique(na.omit(vals))
        observed_vals_str <- if (length(uv) > 8)
          paste(c(head(uv, 8), "..."), collapse = ",")
        else paste(uv, collapse = ",")
      }
    }

    # ---- Format invalid message ----
    invalid_vals_str <- if (has_valid_values) {
      if (all(is.na(invalid_vals)) || length(invalid_vals) == 0)
        "Observed values all valid!"
      else paste(sort(invalid_vals), collapse = ",")
    } else {
      "Need manual check"
    }

    # ================================================================
    # 7. Collect results
    # ================================================================
    results[[varname]] <- data.frame(
      Variable = varname,
      Observed_type = obs_type,
      Expected_type = exp_type,
      Observed_values = observed_vals_str,
      Expected_values = expected_vals,
      Invalid_values = invalid_vals_str,
      stringsAsFactors = FALSE
    )
  }
  
  # ================================================================
  # 8. Combine and return outputs
  # ================================================================
  validation_df <- as.data.frame(do.call(rbind, results))
  corrected_df <- as.data.frame(df)

  cat("Validation complete for dataset:", dataset_name, "\n")

  return(list(
    validation_report = validation_df,
    corrected_df = df
  ))
}

```


<br>

## 1: Subjects

```{r Subjects-1}
df <- as.data.frame(Subjects)

info(df,"e165_ID")

str(df)
```


```{r Subjects-2}
result_df <- validate_dataset("Subjects", IDcol = "e165_ID")

# Validation report
DT::datatable(result_df$validation_report)

# Corrected dataset (with types converted)
Subjects <- as.data.frame(result_df$corrected_df)
str(Subjects)

```


<br>

## 2: Biennial_Visits

```{r Biennial_Visits-1}
df <- as.data.frame(Biennial_Visits)

info(df,"e165_ID")

str(df)
```

```{r Biennial_Visits-2}
result_df <- validate_dataset("Biennial_Visits", IDcol = "e165_ID")

# Validation report
DT::datatable(result_df$validation_report)

# Corrected dataset (with types converted)
Biennial_Visits <- as.data.frame(result_df$corrected_df)
str(Biennial_Visits)

```


<br>

## 3: Form73_Form23_CESD

<br>

## 4: Form75_SexGender

<br>

## 5: Form75_RaceEthnicity

<br>

## 6: Form75_PersonalEducation

<br>

## 7: Form77_Diabetes

<br>

## 8: Form77_Coronary

<br>

## 9: Form77_OtherCardiovascular

<br>

## 10: Form77_Respiratory

<br>

## 11: Form77_Neurological

<br>

## 12: Form77_OtherDisease

<br>

## 13: Form77_SelfReportMeds

<br>

## 14: Form77_VitaminSupplements

<br>

## 15: Form78_InjuryLOC

<br>

## 16: Form78_Smoking

<br>

## 17: Form78_Alcohol

<br>

## 18: Form85_BP

<br>

## 19: Form85_Anthropometry

<br>

## 20:Form55_Short_BISQ

<br>

## 21: Form84_CASI_MMSE

<br>

## 22: Form69_Short_CASI_MMSE

<br>

## 23: Form51_CogTests

<br>

## 24: Form51_Form84_MELV_Scores

<br>


## 25: Form93_Form02_Dementia

<br>


## 26: Form81_CDR

<br>

## 27: genetics_apoe

<br>

## 28: autopsy_status

<br>

## 29: np_standard

<br>

## 30: nic_cdes

<br>

## 31: nic_am_vols

<br>

## 32: nic_pm_vols

<br>

