---
title: "WRAP Data Prep 2025"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
    css: "style.css"
---

Author: Mengna Zhang

*Last updated on:* `r Sys.Date()`

# Set Path

```{r}
#directory
directory <- "/Users/mengnazhang/Desktop/"
my_directory <- paste0(directory, "ADSP_DataPrep_local/WRAP/Phenotype/2025/Raw/")
```

<br>

# Load Packages
```{r,message=FALSE,warning=FALSE}
require(dplyr)
require(tidyr)
require(stringr)
```

<br>

# Load Helper Scripts
```{r}
source("/Users/mengnazhang/Desktop/ADSP_DataPrep/dataPrep2025/helperScripts.R")
```


<br>
<br>

# Variable Check Per File

## Abnormal_Factor_Flags_long

Indicator variables indicating low scores on FactorScores factors, adjusting for demographics.

```{r}
## load data
Abnormal_Factor_Flags_long <- read.csv(paste0(my_directory, "Abnormal_Factor_Flags_long_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Abnormal_Factor_Flags_long,"wrapnum")
```

```{r, eval=FALSE}
str(Abnormal_Factor_Flags_long)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Abnormal_Factor_Flags_long)
```
</details> 

```{r}
## check table frequency for all variables

## drop update_stamp column
Abnormal_Factor_Flags_long <- subset(Abnormal_Factor_Flags_long,select = -c(update_stamp)) ## 6 columns left
```

```{r,eval=FALSE}
str(Abnormal_Factor_Flags_long)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Abnormal_Factor_Flags_long)
```
</details> 


<br>
<br>

## Abnormal_Factor_Flags_wide

Indicator variables indicating low scores on FactorScores factors, adjusting for demographics.

```{r}
## load data
Abnormal_Factor_Flags_wide <- read.csv(paste0(my_directory, "Abnormal_Factor_Flags_wide_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Abnormal_Factor_Flags_wide,"wrapnum")
```

```{r,eval=FALSE}
str(Abnormal_Factor_Flags_wide)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Abnormal_Factor_Flags_wide)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
Abnormal_Factor_Flags_wide <- subset(Abnormal_Factor_Flags_wide,select = -c(update_stamp))
```

```{r,eval=FALSE}
str(Abnormal_Factor_Flags_wide)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Abnormal_Factor_Flags_wide)
```
</details> 


<br>
<br>

## AnimalNaming_ItemLevel

Animal Naming Item Level Data

```{r}
## load data
AnimalNaming_ItemLevel <- read.csv(paste0(my_directory, "AnimalNaming_ItemLevel_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(AnimalNaming_ItemLevel,"wrapnum")
```


```{r,eval=FALSE}
str(AnimalNaming_ItemLevel)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(AnimalNaming_ItemLevel)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(AnimalNaming_ItemLevel)[sapply(AnimalNaming_ItemLevel, is.character)]
AnimalNaming_ItemLevel[chrcols] <- lapply(AnimalNaming_ItemLevel[chrcols], function(x) ifelse(x == "", NA, x))

## convert from logi to character for columns: AN_45 - AN_50
## these columns were confirmed to only have NA values in the raw data
logicols <- paste0("AN_",seq(45,50))
AnimalNaming_ItemLevel[logicols] <- lapply(AnimalNaming_ItemLevel[logicols], as.character)

## drop update_stamp column
AnimalNaming_ItemLevel <- subset(AnimalNaming_ItemLevel,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(AnimalNaming_ItemLevel)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(AnimalNaming_ItemLevel)
```
</details> 


<br>
<br>

## APG

APOE genotype

```{r}
## load data
APG <- read.csv(paste0(my_directory, "APG_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(APG,"wrapnum")
```


```{r,eval=FALSE}
str(APG)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(APG)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
APG <- subset(APG,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(APG)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(APG)
```
</details> 

<br>
<br>

## AVLTPositionScores

Serial position of answers given per trial on Rey AVLT

```{r}
## load data
AVLTPositionScores <- read.csv(paste0(my_directory, "AVLTPositionScores_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(AVLTPositionScores,"wrapnum")
```

```{r,eval=FALSE}
str(AVLTPositionScores, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(AVLTPositionScores, max.level = 99, list.len = 99999)
```
</details> 

```{r}

## convert column B.15 from logistic to integer
## this column was confirmed to only have NA values in the raw data
AVLTPositionScores[,"B.15"] <- as.integer(AVLTPositionScores[,"B.15"])

## check table frequency for all variables

## values in columns ["T1.1" - "D.15"] should be 1-15 (integer)
cols = c(paste0("T1.",seq(1:15)),paste0("T2.",seq(1:15)),paste0("T3.",seq(1:15)),
         paste0("T4.",seq(1:15)),paste0("T5.",seq(1:15)),paste0("B.",seq(1:15)),
         paste0("SD.",seq(1:15)),paste0("D.",seq(1:15)))

for (col in cols) {
  vals <- unique(AVLTPositionScores[[col]])
  invalid_vals <- vals[!vals %in% c(1:15, NA)]
  
  if (length(invalid_vals) > 0) {
    cat("Column:", col, "- Invalid values:", paste(invalid_vals, collapse = ", "), "\n")
  }
}

## drop update_stamp column
AVLTPositionScores <- subset(AVLTPositionScores,select = -c(update_stamp)) ## 124 columns left
```

```{r,eval=FALSE}
str(AVLTPositionScores, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(AVLTPositionScores, max.level = 99, list.len = 99999)
```
</details> 


<br>
<br>

## BehavioralObservationChecklist

Behavioral Observation Checklist

```{r}
## load data
BehavioralObservationChecklist <- read.csv(paste0(my_directory, "BehavioralObservationChecklist_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()

info(BehavioralObservationChecklist,"wrapnum")
```

```{r,eval=FALSE}
str(BehavioralObservationChecklist, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(BehavioralObservationChecklist, max.level = 99, list.len = 99999)
```
</details> 


```{r}
## check table frequency for all variables

## values in columns [BOC_Rapport_EasilyEstab - BOC_Address_Not_PM] should be {1,2}
## 1=selected;2=other value selected;null=no values in set selected
cols = names(BehavioralObservationChecklist)[5:84]

for (col in cols) {
  vals <- unique(BehavioralObservationChecklist[[col]])
  invalid_vals <- vals[!vals %in% c(1:2, NA)]
  
  if (length(invalid_vals) > 0) {
    cat("Column:", col, "- Invalid values:", paste(invalid_vals, collapse = ", "), "\n")
  }
}
## Column: BOC_Motor_Handedness - Invalid values: 3 

table(BehavioralObservationChecklist$BOC_Motor_Handedness, useNA = "ifany")
#    1    2    3 <NA> 
# 2020  214    2  365 

## covert value 3 to NA for column BOC_Motor_Handedness
BehavioralObservationChecklist$BOC_Motor_Handedness[BehavioralObservationChecklist$BOC_Motor_Handedness == 3] <- NA
table(BehavioralObservationChecklist$BOC_Motor_Handedness, useNA = "ifany")
#    1    2 <NA> 
# 2020  214  367 

## drop update_stamp column
BehavioralObservationChecklist <- subset(BehavioralObservationChecklist,select = -c(update_stamp)) ## 84 columns left
```

```{r,eval=FALSE}
str(BehavioralObservationChecklist, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(BehavioralObservationChecklist, max.level = 99, list.len = 99999)
```
</details> 


<br>
<br>

## CA_Genome

TOMM40 genotype

```{r}
## load data
CA_Genome <- read.csv(paste0(my_directory, "CA_Genome_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CA_Genome,"wrapnum")
```

```{r,eval=FALSE}
str(CA_Genome)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CA_Genome)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
CA_Genome <- subset(CA_Genome,select = -c(update_stamp))
```

```{r,eval=FALSE}
str(CA_Genome)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CA_Genome)
```
</details> 


<br>
<br>

## CDR

Clinical Dementia Rating

```{r}
## load data
CDR <- read.csv(paste0(my_directory, "CDR_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CDR,"wrapnum")
```

```{r,eval=FALSE}
str(CDR)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CDR)
```
</details> 


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding. 
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
CDR$RelLive[CDR$RelLive==0] <- 2
CDR$RelLive <- as.integer(CDR$RelLive)

## drop update_stamp column
CDR <- subset(CDR,select = -c(update_stamp)) ## 15 columns left
```

```{r,eval=FALSE}
str(CDR)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CDR)
```
</details> 

<br>
<br>

## CERAD

Consortium to Establish a Registry for Alzheimer's Disease Data (CERAD)

```{r}
## load data
CERAD <- read.csv(paste0(my_directory, "CERAD_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CERAD,"wrapnum")
```



```{r,eval=FALSE}
str(CERAD)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CERAD)
```
</details> 


```{r}
## convert "" to NA for column OverImp
CERAD$OverImp[CERAD$OverImp == ""] <- NA

## drop update_stamp column
CERAD <- subset(CERAD,select = -c(update_stamp)) ## 19 columns left
```

```{r,eval=FALSE}
str(CERAD)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CERAD)
```
</details> 


<br>
<br>

## CES_D_Scale

Center for Epidemiologic Studies-Depression Scale

```{r}
## load data
CES_D_Scale <- read.csv(paste0(my_directory, "CES_D_Scale_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CES_D_Scale,"wrapnum")
```

```{r,eval=FALSE}
str(CES_D_Scale)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CES_D_Scale)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
CES_D_Scale <- subset(CES_D_Scale,select = -c(update_stamp)) ## 24 columns left
```


```{r,eval=FALSE}
str(CES_D_Scale)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CES_D_Scale)
```
</details> 


<br>
<br>

## CHAMPS

Community Healthy Activities Model Program for Seniors (CHAMPS)

```{r}
## load data
CHAMPS <- read.csv(paste0(my_directory, "CHAMPS_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CHAMPS,"wrapnum")
```


```{r,eval=FALSE}
str(CHAMPS, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CHAMPS, max.level = 99, list.len = 99999)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
CHAMPS <- subset(CHAMPS,select = -c(update_stamp)) ## 126 columns left
```


```{r,eval=FALSE}
str(CHAMPS, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CHAMPS, max.level = 99, list.len = 99999)
```
</details> 



<br>
<br>

## CogStateData

Cogstate Testing Data

```{r}
## load data
CogStateData <- read.csv(paste0(my_directory, "CogStateData_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CogStateData,"wrapnum")
```


```{r,eval=FALSE}
str(CogStateData, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CogStateData, max.level = 99, list.len = 99999)
```
</details> 



```{r}
## check table frequency for all variables

## convert logi columns to int
## these columns were confirmed to only have NA values in the raw data
logicols <- colnames(CogStateData)[sapply(CogStateData, is.logical)]
CogStateData[logicols] <- lapply(CogStateData[logicols], as.integer)

## drop update_stamp column
CogStateData <- subset(CogStateData,select = -c(update_stamp)) ## 227 columns left
```


```{r,eval=FALSE}
str(CogStateData, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CogStateData, max.level = 99, list.len = 99999)
```
</details> 



<br>
<br>

## CompositeScoreCalcValues

Raw scores used in computation of CompositeScores values

```{r}
## load data
CompositeScoreCalcValues <- read.csv(paste0(my_directory, "CompositeScoreCalcValues_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CompositeScoreCalcValues,"wrapnum")
```


```{r,eval=FALSE}
str(CompositeScoreCalcValues)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CompositeScoreCalcValues)
```
</details> 



```{r}
## check table frequency for all variables

## drop update_stamp column
CompositeScoreCalcValues <- subset(CompositeScoreCalcValues,select = -c(update_stamp)) ## 16 columns left
```

```{r,eval=FALSE}
str(CompositeScoreCalcValues)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CompositeScoreCalcValues)
```
</details> 


<br>
<br>

## CompositeScores

Domain-specific and global cognitive composite scores (see Jonaitis et al 2019)

```{r}
## load data
CompositeScores <- read.csv(paste0(my_directory, "CompositeScores_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(CompositeScores,"wrapnum")
```


```{r,eval=FALSE}
str(CompositeScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CompositeScores)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
CompositeScores <- subset(CompositeScores,select = -c(update_stamp)) ## 31 columns left
```


```{r,eval=FALSE}
str(CompositeScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(CompositeScores)
```
</details> 


<br>
<br>

## ConsensusConference

Consensus Conference Results

```{r}
## load data
ConsensusConference <- read.csv(paste0(my_directory, "ConsensusConference_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(ConsensusConference,"wrapnum")
```


```{r,eval=FALSE}
str(ConsensusConference)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(ConsensusConference)
```
</details> 


```{r}
## check table frequency for all variables

## convert logi columns to int
## these columns were confirmed to only have NA values in the raw data
logicols <- colnames(ConsensusConference)[sapply(ConsensusConference, is.logical)]
ConsensusConference[logicols] <- lapply(ConsensusConference[logicols], as.integer)

## drop update_stamp column
ConsensusConference <- subset(ConsensusConference,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(ConsensusConference)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(ConsensusConference)
```
</details> 


<br>
<br>

## DailyActivity

Legacy - Daily Activities (Prior to inclusion in General Health and Daily Activities)

```{r}
## load data
DailyActivity <- read.csv(paste0(my_directory, "DailyActivity_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(DailyActivity,"wrapnum")
```


```{r,eval=FALSE}
str(DailyActivity)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(DailyActivity)
```
</details> 


```{r}
## check table frequency for all variables

## convert logi columns to int
## ## these columns were confirmed to only have NA values in the raw data
logicols <- colnames(DailyActivity)[sapply(DailyActivity, is.logical)]
DailyActivity[logicols] <- lapply(DailyActivity[logicols], as.integer)

## extract all chr columns and recode "" to NA
chrcols <- colnames(DailyActivity)[sapply(DailyActivity, is.character)]
DailyActivity[chrcols] <- lapply(DailyActivity[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
DailyActivity <- subset(DailyActivity,select = -c(update_stamp)) ## 84 columns left
```


```{r,eval=FALSE}
str(DailyActivity)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(DailyActivity)
```
</details> 


<br>
<br>

## Demographics

Participant Demographic Information

```{r}
## load data
Demographics <- read.csv(paste0(my_directory, "Demographics_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Demographics,"wrapnum")
```


```{r,eval=FALSE}
str(Demographics)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Demographics)
```
</details> 


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding.
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
Demographics$OC_Imputed_grade[Demographics$OC_Imputed_grade] <- 2
Demographics$OC_Imputed_grade <- as.integer(Demographics$OC_Imputed_grade)

## convert logi columns to int
logicols <- colnames(Demographics)[sapply(Demographics, is.logical)]
Demographics[logicols] <- lapply(Demographics[logicols], as.integer)

## convert parntres from num to int
Demographics$parntres <- as.integer(Demographics$parntres)

## convert "" to NA for chr columns
chrcols <- colnames(Demographics)[sapply(Demographics, is.character)]
Demographics[chrcols] <- lapply(Demographics[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Demographics <- subset(Demographics,select = -c(update_stamp)) ## 84 columns left
```


```{r,eval=FALSE}
str(Demographics)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Demographics)
```
</details> 

<br>
<br>

## FactorScores

Factor scores based on analysis of Wave 1 and Wave 2 neuropsychological test data (see Dowling et al. 2010).

```{r}
## load data
FactorScores <- read.csv(paste0(my_directory, "FactorScores_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(FactorScores,"wrapnum")
```


```{r,eval=FALSE}
str(FactorScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(FactorScores)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
FactorScores <- subset(FactorScores,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(FactorScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(FactorScores)
```
</details> 


<br>
<br>

## fqryStatisticalData

Participant/Visit-level study data

```{r}
## load data
fqryStatisticalData <- read.csv(paste0(my_directory, "fqryStatisticalData_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(fqryStatisticalData,"wrapnum")
```


```{r,eval=FALSE}
str(fqryStatisticalData)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(fqryStatisticalData)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(fqryStatisticalData)[sapply(fqryStatisticalData, is.character)]
fqryStatisticalData[chrcols] <- lapply(fqryStatisticalData[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
fqryStatisticalData <- subset(fqryStatisticalData,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(fqryStatisticalData)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(fqryStatisticalData)
```
</details> 


<br>
<br>

## Gene_DataCollection

Gene Data Collection

```{r}
## load data
Gene_DataCollection <- read.csv(paste0(my_directory, "Gene_DataCollection_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Gene_DataCollection,"wrapnum")
```


```{r,eval=FALSE}
str(Gene_DataCollection)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Gene_DataCollection)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(Gene_DataCollection)[sapply(Gene_DataCollection, is.character)]
Gene_DataCollection[chrcols] <- lapply(Gene_DataCollection[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Gene_DataCollection <- subset(Gene_DataCollection,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Gene_DataCollection)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Gene_DataCollection)
```
</details> 


<br>
<br>

## GenHealthDailyAct

General Health and Daily Activities

```{r}
## load data
GenHealthDailyAct <- read.csv(paste0(my_directory, "GenHealthDailyAct_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(GenHealthDailyAct,"wrapnum")
```


```{r,eval=FALSE}
str(GenHealthDailyAct,max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(GenHealthDailyAct,max.level = 99, list.len = 99999)
```
</details> 


```{r}
## check table frequency for all variables

## convert 6 to NA for columns: YardHrs,FTFFAMPL,FTFMEETINSSPL,FTFSTRANGERPL
tmp <- c("YardHrs","FTFFAMPL","FTFMEETINSSPL","FTFSTRANGERPL")
GenHealthDailyAct[tmp] <- lapply(GenHealthDailyAct[tmp], function(x) {
  x[x == 6] <- NA
  return(x)
})

## convert 5 to NA for column: FTFOTHER
GenHealthDailyAct$FTFOTHER[GenHealthDailyAct$FTFOTHER==5] <- NA

## convert 8 to NA for column: FTFSPPARTPL
GenHealthDailyAct$FTFSPPARTPL[GenHealthDailyAct$FTFSPPARTPL==8] <- NA

## convert "" to NA for chr columns
chrcols <- colnames(GenHealthDailyAct)[sapply(GenHealthDailyAct, is.character)]
GenHealthDailyAct[chrcols] <- lapply(GenHealthDailyAct[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
GenHealthDailyAct <- subset(GenHealthDailyAct,select = -c(update_stamp)) ## 219 columns left
```


```{r,eval=FALSE}
str(GenHealthDailyAct,max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(GenHealthDailyAct,max.level = 99, list.len = 99999)
```
</details> 


<br>
<br>

## HeadInjuries

Head Injuries

```{r}
## load data
HeadInjuries <- read.csv(paste0(my_directory, "HeadInjuries_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(HeadInjuries,"wrapnum")
```


```{r,eval=FALSE}
str(HeadInjuries)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(HeadInjuries)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
HeadInjuries <- subset(HeadInjuries,select = -c(update_stamp)) ## 4 columns left
```


```{r,eval=FALSE}
str(HeadInjuries)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(HeadInjuries)
```
</details> 

<br>
<br>

## HeadInjurySub

Head Injuries - Additional Data

```{r}
## load data
HeadInjurySub <- read.csv(paste0(my_directory, "HeadInjurySub_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(HeadInjurySub,"wrapnum")
```


```{r,eval=FALSE}
str(HeadInjurySub)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(HeadInjurySub)
```
</details> 

```{r}
## check table frequency for all variables

## drop update_stamp column
HeadInjurySub <- subset(HeadInjurySub,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(HeadInjurySub)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(HeadInjurySub)
```
</details> 


<br>
<br>

## IADL

Lawton Instrumental Activities of Daily Living (IADL)

```{r}
## load data
IADL <- read.csv(paste0(my_directory, "IADL_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(IADL,"wrapnum")
```


```{r,eval=FALSE}
str(IADL)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(IADL)
```
</details> 


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding. 
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
IADL$RelLive[IADL$RelLive==0] <- 2
IADL$RelLive <- as.integer(IADL$RelLive)

## drop update_stamp column
IADL <- subset(IADL,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(IADL)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(IADL)
```
</details> 


<br>
<br>

## IQCode

```{r}
## load data
IQCode <- read.csv(paste0(my_directory, "IQCode_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(IQCode,"wrapnum")
```


```{r,eval=FALSE}
str(IQCode)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(IQCode)
```
</details> 


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding. 
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
IQCode$RelLive[IQCode$RelLive==0] <- 2
IQCode$RelLive <- as.integer(IQCode$RelLive)

## drop update_stamp column
IQCode <- subset(IQCode,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(IQCode)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(IQCode)
```
</details> 


<br>
<br>

## LearningDifficulties

Learning Difficulties interview

```{r}
## load data
LearningDifficulties <- read.csv(paste0(my_directory, "LearningDifficulties_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(LearningDifficulties,"wrapnum")
```


```{r,eval=FALSE}
str(LearningDifficulties)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(LearningDifficulties)
```
</details> 


```{r}
## check table frequency for all variables

## convert logi columns to int
## these columns were confirmed to only have NA values in the raw data
logicols <- colnames(LearningDifficulties)[sapply(LearningDifficulties, is.logical)]
LearningDifficulties[logicols] <- lapply(LearningDifficulties[logicols], as.integer)

## convert "" to NA for chr columns
chrcols <- colnames(LearningDifficulties)[sapply(LearningDifficulties, is.character)]
LearningDifficulties[chrcols] <- lapply(LearningDifficulties[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
LearningDifficulties <- subset(LearningDifficulties,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(LearningDifficulties)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(LearningDifficulties)
```
</details> 


<br>
<br>

## LIBRA

Lifestyle for BRAin health

```{r}
## load data
LIBRA <- read.csv(paste0(my_directory, "LIBRA_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(LIBRA,"wrapnum")
```


```{r,eval=FALSE}
str(LIBRA)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(LIBRA)
```
</details> 


```{r}
## check table frequency for all variables

## drop update_stamp column
LIBRA <- subset(LIBRA,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(LIBRA)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(LIBRA)
```
</details> 


<br>
<br>

## Med_Lab

Med/Lab Scores

```{r}
## load data
Med_Lab <- read.csv(paste0(my_directory, "Med_Lab_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Med_Lab,"wrapnum")
```


```{r,eval=FALSE}
str(Med_Lab)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab)
```
</details> 


```{r}
## check table frequency for all variables

## DD has 1=Yes;0=No;3=Unknown coding but data has 0/1/2 coding. 
## I will convert 2 to 3 in the dataset
Med_Lab$antipsychotic_reaction[Med_Lab$antipsychotic_reaction==2] <- 3
Med_Lab$antipsychotic_reaction <- as.integer(Med_Lab$antipsychotic_reaction)

## convert "" to NA for chr columns
chrcols <- colnames(Med_Lab)[sapply(Med_Lab, is.character)]
Med_Lab[chrcols] <- lapply(Med_Lab[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Med_Lab <- subset(Med_Lab,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Med_Lab)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab)
```
</details> 


<br>
<br>

## Med_Lab_NPDC

Med/Lab Non-Prescription Drug Therapy

```{r}
## load data
Med_Lab_NPDC <- read.csv(paste0(my_directory, "Med_Lab_NPDC_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Med_Lab_NPDC,"wrapnum")
```


```{r,eval=FALSE}
str(Med_Lab_NPDC)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab_NPDC)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(Med_Lab_NPDC)[sapply(Med_Lab_NPDC, is.character)]
Med_Lab_NPDC[chrcols] <- lapply(Med_Lab_NPDC[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Med_Lab_NPDC <- subset(Med_Lab_NPDC,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Med_Lab_NPDC)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab_NPDC)
```
</details> 


<br>
<br>

## Med_Lab_PDC

Med/Lab Prescription Drug Therapy

```{r}
## load data
Med_Lab_PDC <- read.csv(paste0(my_directory, "Med_Lab_PDC_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Med_Lab_PDC,"wrapnum")
```


```{r,eval=FALSE}
str(Med_Lab_PDC)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab_PDC)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(Med_Lab_PDC)[sapply(Med_Lab_PDC, is.character)]
Med_Lab_PDC[chrcols] <- lapply(Med_Lab_PDC[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Med_Lab_PDC <- subset(Med_Lab_PDC,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Med_Lab_PDC)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Med_Lab_PDC)
```
</details> 


<br>
<br>

## MedHistory

Medical History

```{r}
## load data
MedHistory <- read.csv(paste0(my_directory, "MedHistory_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MedHistory,"wrapnum")
```


```{r,eval=FALSE}
str(MedHistory)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory)
```
</details> 


```{r}
## check table frequency for all variables

## values in columns [AdmitS,SurgS - HeadInj_E, Unc_E, Stroke_E - Neuro_E, Sleep_E - TumorOrCancer_E, seehear] should be in {1,2,3}
cols = c("AdmitS",
         names(MedHistory)[which(names(MedHistory) == "SurgS") : which(names(MedHistory) == "HeadInj_E")],
         "Unc_E",
         names(MedHistory)[which(names(MedHistory) == "Stroke_E") : which(names(MedHistory) == "Neuro_E")],
         names(MedHistory)[which(names(MedHistory) == "Sleep_E") : which(names(MedHistory) == "TumorOrCancer_E")],
         "seehear"
        )

for (col in cols) {
  vals <- unique(MedHistory[[col]])
  invalid_vals <- vals[!vals %in% c(1,2,3, NA)]
  
  if (length(invalid_vals) > 0) {
    cat("Column:", col, "- Invalid values:", paste(invalid_vals, collapse = ", "), "\n")
  }
}

## drop update_stamp column
MedHistory <- subset(MedHistory,select = -c(update_stamp))
```

```{r,eval=FALSE}
str(MedHistory)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory)
```
</details> 


<br>
<br>

## Morbidity_Mortality

Participant Morbidity and Mortality Details

```{r}
## load data
Morbidity_Mortality <- read.csv(paste0(my_directory, "Morbidity_Mortality_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Morbidity_Mortality,"wrapnum")
```


```{r,eval=FALSE}
str(Morbidity_Mortality)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Morbidity_Mortality)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(Morbidity_Mortality)[sapply(Morbidity_Mortality, is.character)]
Morbidity_Mortality[chrcols] <- lapply(Morbidity_Mortality[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Morbidity_Mortality <- subset(Morbidity_Mortality,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Morbidity_Mortality)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Morbidity_Mortality)
```
</details> 


<br>
<br>

## MedHistory_Admit

Medical History (Admitted to Hospital)

```{r}
## load data
MedHistory_Admit <- read.csv(paste0(my_directory, "MedHistory_Admit_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MedHistory_Admit,"wrapnum")
```


```{r,eval=FALSE}
str(MedHistory_Admit)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_Admit)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(MedHistory_Admit)[sapply(MedHistory_Admit, is.character)]
MedHistory_Admit[chrcols] <- lapply(MedHistory_Admit[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
MedHistory_Admit <- subset(MedHistory_Admit,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MedHistory_Admit)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_Admit)
```
</details> 


<br>
<br>

## MedHistory_SurgBF

Medical History (Surgery since last visit)

```{r}
## load data
MedHistory_SurgBF <- read.csv(paste0(my_directory, "MedHistory_SurgBF_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MedHistory_SurgBF,"wrapnum")
```


```{r,eval=FALSE}
str(MedHistory_SurgBF)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_SurgBF)
```
</details> 


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(MedHistory_SurgBF)[sapply(MedHistory_SurgBF, is.character)]
MedHistory_SurgBF[chrcols] <- lapply(MedHistory_SurgBF[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
MedHistory_SurgBF <- subset(MedHistory_SurgBF,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MedHistory_SurgBF)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_SurgBF)
```


<br>
<br>

## MedHistory_SurgS

Medical History (Surgery since last visit)

```{r}
## load data
MedHistory_SurgS <- read.csv(paste0(my_directory, "MedHistory_SurgS_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MedHistory_SurgS,"wrapnum")
```


```{r,eval=FALSE}
str(MedHistory_SurgS)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_SurgS)
```


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(MedHistory_SurgS)[sapply(MedHistory_SurgS, is.character)]
MedHistory_SurgS[chrcols] <- lapply(MedHistory_SurgS[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
MedHistory_SurgS <- subset(MedHistory_SurgS,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MedHistory_SurgS)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MedHistory_SurgS)
```


<br>
<br>

## MIND_Diet

MIND Diet

```{r}
## load data
MIND_Diet <- read.csv(paste0(my_directory, "MIND_Diet_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MIND_Diet,"wrapnum")
```


```{r,eval=FALSE}
str(MIND_Diet)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MIND_Diet)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
MIND_Diet <- subset(MIND_Diet,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MIND_Diet)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MIND_Diet)
```



<br>
<br>

## MIND_Diet_ScoredValues

MIND Diet Scored Values

This dataset is exactly same as the NIND_Diet dataset. So I won't be including this file when do the final merge
```{r}
## load data
MIND_Diet_ScoredValues <- read.csv(paste0(my_directory, "MIND_Diet_ScoredValues_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MIND_Diet_ScoredValues,"wrapnum")
```


```{r,eval=FALSE}
str(MIND_Diet_ScoredValues)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MIND_Diet_ScoredValues)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
MIND_Diet <- subset(MIND_Diet_ScoredValues,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MIND_Diet_ScoredValues)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MIND_Diet_ScoredValues)
```



<br>
<br>

## MOCA

Montreal Cognitive Assessment-Blind. Administered as part of remote assessment (T-Cog)

```{r}
## load data
MOCA <- read.csv(paste0(my_directory, "MOCA_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(MOCA,"wrapnum")
```


```{r,eval=FALSE}
str(MOCA)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MOCA)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
MOCA <- subset(MOCA,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(MOCA)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(MOCA)
```


<br>
<br>

## NeuropsychScores

Neuropsych Testing Scores

```{r}
## load data
NeuropsychScores <- read.csv(paste0(my_directory, "NeuropsychScores_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(NeuropsychScores,"wrapnum")
```


```{r,eval=FALSE}
str(NeuropsychScores, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(NeuropsychScores, max.level = 99, list.len = 99999)
```


```{r}
## convert "" to NA for chr columns
chrcols <- colnames(NeuropsychScores)[sapply(NeuropsychScores, is.character)]
NeuropsychScores[chrcols] <- lapply(NeuropsychScores[chrcols], function(x) ifelse(x == "", NA, x))

## convert from logi to int for columns: wmfacpr wmfacps animPhoMean animPhoSwi cflTotRaw cflSemMean cflSemSwi
## these columns were confirmed to only have NA values in the raw data
logicols <- c("wmfacpr","wmfacps","animPhoMean","animPhoSwi","cflTotRaw","cflSemMean","cflSemSwi")
NeuropsychScores[logicols] <- lapply(NeuropsychScores[logicols], as.integer)

## drop update_stamp column
NeuropsychScores <- subset(NeuropsychScores,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(NeuropsychScores, max.level = 99, list.len = 99999)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(NeuropsychScores, max.level = 99, list.len = 99999)
```


<br>
<br>

## PersFamMedBack

Personal Family Medical Background

```{r}
## load data
PersFamMedBack <- read.csv(paste0(my_directory, "PersFamMedBack_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(PersFamMedBack,"wrapnum")
```


```{r,eval=FALSE}
str(PersFamMedBack)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(PersFamMedBack)
```


```{r}
## check table frequency for all variables

## for column PersFamMedBack, in DD it got values:1=Yes;0=No
## but in the dataset, it has value:
# table(PersFamMedBack$Volunt,useNA = "ifany")
# 
#    0    1    2 <NA> 
#  114 1935 1530 4597
## by comparing with how other value being coded, I decide to convert 2 to 0
PersFamMedBack$Volunt[PersFamMedBack$Volunt==2] <- 0

## convert "" to NA for chr columns
chrcols <- colnames(PersFamMedBack)[sapply(PersFamMedBack, is.character)]
PersFamMedBack[chrcols] <- lapply(PersFamMedBack[chrcols], function(x) ifelse(x == "", NA, x))

## convert logi to chr for column: gender_other,gender_identity_other
logicols <- c("gender_other","gender_identity_other")
PersFamMedBack[logicols] <- lapply(PersFamMedBack[logicols], as.character)

## drop update_stamp column
PersFamMedBack <- subset(PersFamMedBack,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(PersFamMedBack)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(PersFamMedBack)
```


<br>
<br>


## PFMB_SibWAD_or_OtherMemIssue

Personal Family Medical Background - Siblings with AD or Other Memory Issue

```{r}
## load data
PFMB_SibWAD_or_OtherMemIssue <- read.csv(paste0(my_directory, "PFMB_SibWAD_or_OtherMemIssue_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(PFMB_SibWAD_or_OtherMemIssue,"wrapnum")
```


```{r,eval=FALSE}
str(PFMB_SibWAD_or_OtherMemIssue)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(PFMB_SibWAD_or_OtherMemIssue)
```


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(PFMB_SibWAD_or_OtherMemIssue)[sapply(PFMB_SibWAD_or_OtherMemIssue, is.character)]
PFMB_SibWAD_or_OtherMemIssue[chrcols] <- lapply(PFMB_SibWAD_or_OtherMemIssue[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
PFMB_SibWAD_or_OtherMemIssue <- subset(PFMB_SibWAD_or_OtherMemIssue,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(PFMB_SibWAD_or_OtherMemIssue)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(PFMB_SibWAD_or_OtherMemIssue)
```


<br>
<br>


## QuickDementia

Quick Dementia Rating Scale

```{r}
## load data
QuickDementia <- read.csv(paste0(my_directory, "QuickDementia_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(QuickDementia,"wrapnum")
```


```{r,eval=FALSE}
str(QuickDementia)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(QuickDementia)
```


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding. 
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
QuickDementia$RelLive[QuickDementia$RelLive==0] <- 2
QuickDementia$RelLive <- as.integer(QuickDementia$RelLive)

## drop update_stamp column
QuickDementia <- subset(QuickDementia,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(QuickDementia)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(QuickDementia)
```


<br>
<br>


## RobustNormZScores

Demographically-adjusted robust norm z scores (see Koscik et al. 2014, Clark et al. 2016)

```{r}
## load data
RobustNormZScores <- read.csv(paste0(my_directory, "RobustNormZScores_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(RobustNormZScores,"wrapnum")
```


```{r,eval=FALSE}
str(RobustNormZScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(RobustNormZScores)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
RobustNormZScores <- subset(RobustNormZScores,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(RobustNormZScores)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(RobustNormZScores)
```


<br>
<br>


## Sleep_MentalAct

Sleep/Mental Activities

```{r}
## load data
Sleep_MentalAct <- read.csv(paste0(my_directory, "Sleep_MentalAct_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Sleep_MentalAct,"wrapnum")
```


```{r,eval=FALSE}
str(Sleep_MentalAct)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Sleep_MentalAct)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
Sleep_MentalAct <- subset(Sleep_MentalAct,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Sleep_MentalAct)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Sleep_MentalAct)
```


<br>
<br>


## Sleep_MentalAct_OthDevice

Sleep/Mental Activities - Sleep Apnea Other Devices

```{r}
## load data
Sleep_MentalAct_OthDevice <- read.csv(paste0(my_directory, "Sleep_MentalAct_OthDevice_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Sleep_MentalAct_OthDevice,"wrapnum")
```


```{r,eval=FALSE}
str(Sleep_MentalAct_OthDevice)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Sleep_MentalAct_OthDevice)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
Sleep_MentalAct_OthDevice <- subset(Sleep_MentalAct_OthDevice,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(Sleep_MentalAct_OthDevice)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Sleep_MentalAct_OthDevice)
```

<br>
<br>

## SubjectiveHearing

Subjective Hearing Complaints

```{r}
## load data
SubjectiveHearing <- read.csv(paste0(my_directory, "SubjectiveHearing_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(SubjectiveHearing,"wrapnum")
```


```{r,eval=FALSE}
str(SubjectiveHearing)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(SubjectiveHearing)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
SubjectiveHearing <- subset(SubjectiveHearing,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(SubjectiveHearing)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(SubjectiveHearing)
```


<br>
<br>

## TICSM

Modified Telephone Interview of Cognitive Status

```{r}
## load data
TICSM <- read.csv(paste0(my_directory, "TICSM_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(TICSM,"wrapnum")
```


```{r,eval=FALSE}
str(TICSM)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(TICSM)
```


```{r}
## check table frequency for all variables

## drop update_stamp column
TICSM <- subset(TICSM,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(TICSM)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(TICSM)
```

<br>
<br>

## VisitVitals

WRAP Visit Vitals

```{r}
## load data
VisitVitals <- read.csv(paste0(my_directory, "VisitVitals_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(VisitVitals,"wrapnum")
```


```{r,eval=FALSE}
str(VisitVitals)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(VisitVitals)
```

```{r}
## convert "" to NA for chr columns
chrcols <- colnames(VisitVitals)[sapply(VisitVitals, is.character)]
VisitVitals[chrcols] <- lapply(VisitVitals[chrcols], function(x) ifelse(x == "", NA, x))

## convert logi to chr for columns: BP_not_ordered BP_single_measurement
logicols <- c("BP_not_ordered","BP_single_measurement")
VisitVitals[logicols] <- lapply(VisitVitals[logicols], as.character)

## drop update_stamp column
VisitVitals <- subset(VisitVitals,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(VisitVitals)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(VisitVitals)
```

<br>
<br>

## WomenQuestions

Women Questions

```{r}
## load data
WomenQuestions <- read.csv(paste0(my_directory, "WomenQuestions_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions)
```


```{r}
## check table frequency for all variables

## in the DD, column nowhrt have three values: 1=Yes;2=No;3=Don't Know
## in the data, it got addition value 0:
# table(WomenQuestions$nowhrt,useNA="ifany")
# 
#    0    1    2    3 <NA> 
#    1  150  292    3 5098 

## I am going to convert 0 to NA:
WomenQuestions$nowhrt[WomenQuestions$nowhrt==0] <- NA

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions)[sapply(WomenQuestions, is.character)]
WomenQuestions[chrcols] <- lapply(WomenQuestions[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WomenQuestions <- subset(WomenQuestions,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WomenQuestions)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions)
```


<br>
<br>

## WomenQuestions_BCP

Women Questions

```{r}
## load data
WomenQuestions_BCP <- read.csv(paste0(my_directory, "WomenQuestions_BCP_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions_BCP,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions_BCP)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_BCP)
```


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions_BCP)[sapply(WomenQuestions_BCP, is.character)]
WomenQuestions_BCP[chrcols] <- lapply(WomenQuestions_BCP[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WomenQuestions_BCP <- subset(WomenQuestions_BCP,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WomenQuestions_BCP)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_BCP)
```


<br>
<br>

## WomenQuestions_Current_HormoneTherapy

Women Questions - Current Hormone Therapy

```{r}
## load data
WomenQuestions_Current_HormoneTherapy <- read.csv(paste0(my_directory, "WomenQuestions_Current_HormoneTherapy_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions_Current_HormoneTherapy,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions_Current_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_Current_HormoneTherapy)
```


```{r}
## check table frequency for all variables

## DD has 1=yes,2=no coding but data has 0/1 coding. 
## Confirmed with other 0/1 variables that 0 is no and 1 stays yes, so will update the dataset to match the DD
WomenQuestions_Current_HormoneTherapy$CurCur[WomenQuestions_Current_HormoneTherapy$CurCur==0] <- 2
WomenQuestions_Current_HormoneTherapy$CurCur <- as.integer(WomenQuestions_Current_HormoneTherapy$CurCur)

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions_Current_HormoneTherapy)[sapply(WomenQuestions_Current_HormoneTherapy, is.character)]
WomenQuestions_Current_HormoneTherapy[chrcols] <- lapply(WomenQuestions_Current_HormoneTherapy[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WomenQuestions_Current_HormoneTherapy <- subset(WomenQuestions_Current_HormoneTherapy,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WomenQuestions_Current_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_Current_HormoneTherapy)
```



<br>
<br>

## WomenQuestions_NonPresc_HormoneTherapy

Women Questions - Non-Prescription Hormone Therapy
```{r}
## load data
WomenQuestions_NonPresc_HormoneTherapy <- read.csv(paste0(my_directory, "WomenQuestions_NonPresc_HormoneTherapy_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions_NonPresc_HormoneTherapy,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions_NonPresc_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_NonPresc_HormoneTherapy)
```


```{r}
## check table frequency for all variables

## for column HormCur, it has additional value 0 in the data. I did not modify it for now.
table(WomenQuestions_NonPresc_HormoneTherapy$HormCur,useNA = "ifany")

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions_NonPresc_HormoneTherapy)[sapply(WomenQuestions_NonPresc_HormoneTherapy, is.character)]
WomenQuestions_NonPresc_HormoneTherapy[chrcols] <- lapply(WomenQuestions_NonPresc_HormoneTherapy[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WomenQuestions_NonPresc_HormoneTherapy <- subset(WomenQuestions_NonPresc_HormoneTherapy,select = -c(update_stamp))

```


```{r,eval=FALSE}
str(WomenQuestions_NonPresc_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_NonPresc_HormoneTherapy)
```

<br>
<br>

## WomenQuestions_Past_HormoneTherapy

Women Questions - Past Hormone Therapy

```{r}
## load data
WomenQuestions_Past_HormoneTherapy <- read.csv(paste0(my_directory, "WomenQuestions_Past_HormoneTherapy_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions_Past_HormoneTherapy,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions_Past_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_Past_HormoneTherapy)
```


```{r}
## check table frequency for all variables

## for column PastCur, it has additional value 0 in the data. I did not modify it for now.
table(WomenQuestions_Past_HormoneTherapy$PastCur,useNA = "ifany")

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions_Past_HormoneTherapy)[sapply(WomenQuestions_Past_HormoneTherapy, is.character)]
WomenQuestions_Past_HormoneTherapy[chrcols] <- lapply(WomenQuestions_Past_HormoneTherapy[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WomenQuestions_Past_HormoneTherapy <- subset(WomenQuestions_Past_HormoneTherapy,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WomenQuestions_Past_HormoneTherapy)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_Past_HormoneTherapy)
```

<br>
<br>

## WomenQuestions_SERM

Women Questions - SERM (Antiestrogen medication)

```{r}
## load data
WomenQuestions_SERM <- read.csv(paste0(my_directory, "WomenQuestions_SERM_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WomenQuestions_SERM,"wrapnum")
```


```{r,eval=FALSE}
str(WomenQuestions_SERM)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_SERM)
```


```{r}
## check table frequency for all variables

## for column SERMCur, it has additional value 0 in the data. I did not modify it for now.
table(WomenQuestions_SERM$SERMCur,useNA = "ifany")

## convert "" to NA for chr columns
chrcols <- colnames(WomenQuestions_SERM)[sapply(WomenQuestions_SERM, is.character)]
WomenQuestions_SERM[chrcols] <- lapply(WomenQuestions_SERM[chrcols], function(x) ifelse(x == "", NA, x))

## convert column SERMType from logi to chr
WomenQuestions_SERM[["SERMType"]] <- as.character(WomenQuestions_SERM[["SERMType"]])

## drop update_stamp column
WomenQuestions_SERM <- subset(WomenQuestions_SERM,select = -c(update_stamp))
```



```{r,eval=FALSE}
str(WomenQuestions_SERM)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WomenQuestions_SERM)
```


<br>
<br>

## WorkHistory

Participant Work History

```{r}
## load data
WorkHistory <- read.csv(paste0(my_directory, "WorkHistory_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WorkHistory,"wrapnum")
```


```{r,eval=FALSE}
str(WorkHistory)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WorkHistory)
```


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(WorkHistory)[sapply(WorkHistory, is.character)]
WorkHistory[chrcols] <- lapply(WorkHistory[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WorkHistory <- subset(WorkHistory,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WorkHistory)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WorkHistory)
```


<br>
<br>

## WorkHistory_Family

Participant Family Work History

```{r}
## load data
WorkHistory_Family <- read.csv(paste0(my_directory, "WorkHistory_Family_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WorkHistory_Family,"wrapnum")
```


```{r,eval=FALSE}
str(WorkHistory_Family)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WorkHistory_Family)
```


```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(WorkHistory_Family)[sapply(WorkHistory_Family, is.character)]
WorkHistory_Family[chrcols] <- lapply(WorkHistory_Family[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
WorkHistory_Family <- subset(WorkHistory_Family,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WorkHistory_Family)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WorkHistory_Family)
```


<br>
<br>

## WRAP_to_DBID

WRAP to DBID Identifier crosswalk table

```{r}
## load data
WRAP_to_DBID <- read.csv(paste0(my_directory, "WRAP_to_DBID_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(WRAP_to_DBID,"wrapnum")
```


```{r,eval=FALSE}
str(WRAP_to_DBID)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WRAP_to_DBID)
```


```{r}
## drop update_stamp column
WRAP_to_DBID <- subset(WRAP_to_DBID,select = -c(update_stamp))
```


```{r,eval=FALSE}
str(WRAP_to_DBID)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(WRAP_to_DBID)
```

<br>
<br>

## Imaging_metadata

WRAP Imaging Metadata

```{r}
## load data
Imaging_metadata <- read.csv(paste0(my_directory, "Imaging_metadata_05Jun2025.csv"),stringsAsFactors = FALSE) %>% as.data.frame()
info(Imaging_metadata,"wrapnum")
```


```{r,eval=FALSE}
str(Imaging_metadata)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Imaging_metadata)
```

```{r}
## check table frequency for all variables

## convert "" to NA for chr columns
chrcols <- colnames(Imaging_metadata)[sapply(Imaging_metadata, is.character)]
Imaging_metadata[chrcols] <- lapply(Imaging_metadata[chrcols], function(x) ifelse(x == "", NA, x))

## drop update_stamp column
Imaging_metadata <- subset(Imaging_metadata,select = -c(update_stamp))
```



```{r,eval=FALSE}
str(Imaging_metadata)
```
<details>
<summary>Click for details</summary>
```{r, echo=FALSE, eval=TRUE}
str(Imaging_metadata)
```


<br>


# Get Sub-Dataset Names
```{r}
df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
```


# Duplicates Handling

## Duplicates Detection

```{r}
# Get names of all data frames in the environment
df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
dfwithDuplicates <- c()

## CODE provided by Alaina
# Loop through each data frame
# for (df_name in df_names) {
#   # Get the actual data frame
#   df <- get(df_name)
#   # Check if columns ID and Visit exist
#   if (all(c("wrapnum", "VisNo") %in% colnames(df))) {
#     # Find duplicates using dplyr
#     dup_rows <- df %>% 
#       dplyr::group_by(wrapnum, VisNo) %>% 
#       dplyr::filter(n() > 1) %>% 
#       dplyr::ungroup()
#     # If any duplicates found, assign to new data frame with _Duplicates
#     if (nrow(dup_rows) > 0) {
#       new_name <- paste0("Duplicates_",df_name)
#       assign(new_name, dup_rows)
#       cat("Duplicates found and saved to:", new_name, "\n")
#     } else {
#       cat("No duplicates found in:", df_name, "\n")
#     }
#   } else {
#     cat("Skipping", df_name, "- ID and/or Visit column not found\n")
#   }
# }

## Modified Code
for (df_name in df_names) {
  # Get the actual data frame
  df <- get(df_name)
  # Check if columns ID and Visit exist
  if (all(c("wrapnum", "VisNo") %in% colnames(df))) {
    # Find duplicates using dplyr
    dup_rows <- df %>% 
      dplyr::group_by(wrapnum, VisNo) %>% 
      dplyr::filter(n() > 1) %>% 
      dplyr::ungroup()
    # If any duplicates found, assign to new data frame with _Duplicates
    if (nrow(dup_rows) > 0) {
      
      dfwithDuplicates <- c(dfwithDuplicates,df_name)
      
    } else {
      cat("No duplicates found in:", df_name, "\n")
    }
  } else {
    cat("Skipping", df_name, "- ID and/or Visit column not found\n")
  }
}
```


check each skipped file, making sure there is no duplicates

```{r}
for (datname in c("Abnormal_Factor_Flags_wide","APG","CA_Genome","CERAD","Demographics","Gene_DataCollection","Imaging_metadata","Morbidity_Mortality","WorkHistory","WorkHistory_Family","WRAP_to_DBID")) {
  
  dat <- get(datname)
  
  cat(datname,": ", length(unique(dat[["wrapnum"]])) == nrow(dat),"\n")
}

```


```{r}
print(dfwithDuplicates)
```

## Duplicates Handling Per Study

###  Imaging_metadata

```{r}
## the imaging_visit is coded inconsistenly across the observations, so I generated a new numeric column for visit index
Imaging_metadata <- Imaging_metadata %>%
  mutate(
    VisNo = str_extract(imaging_visit, "\\d+") %>% as.numeric()
  )

## group by subject ID and visit index, we are going to retain only one obs per group. The selection criteria is as following: 
## 1. pick the obs with least missing values
## 2. if same, then randomly pick one

Imaging_metadata <- Imaging_metadata %>%
  mutate(missing_count = rowSums(is.na(.))) %>%  # Count missing values per row
  group_by(wrapnum, VisNo) %>%
  arrange(missing_count) %>%
  slice_sample(n = 1, weight_by = (missing_count == min(missing_count))) %>%
  ungroup() %>%
  select(-missing_count)  # Remove helper column

info(Imaging_metadata,"wrapnum")
```


<br>

### WorkHistory

```{r}
## check unique values for each columns, for columns with just one unique value,
## we leave it out from the pivot_wider function
result <- lapply(names(WorkHistory), function(x) {
  cat(x, " ", length(unique(WorkHistory[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(WorkHistory),"\n")

tmp <- WorkHistory %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo),
    names_from = work_history_repeat_key,
    values_from = c(JobYrs:OccZone),
    names_sep = "_"
  )
## drop columns if all obs are missing
WorkHistory <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(WorkHistory),"\n")

```

<br>


### HeadInjurySub

```{r}
## check unique values for each columns, for columns with just one unique value,
## we leave it out from the pivot_wider function
result <- lapply(names(HeadInjurySub), function(x) {
  cat(x, " ", length(unique(HeadInjurySub[[x]])), "\n")
})
```


```{r}
cat("Before duplicates handling: ",dim(HeadInjurySub),"\n")

tmp <- HeadInjurySub %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = RowNumber,
    values_from = c(agehi:cognow),
    names_sep = "_"
  )
## drop columns if all obs are missing
HeadInjurySub <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(HeadInjurySub),"\n")

```


<br>

### Med_Lab_NPDC

```{r}
result <- lapply(names(Med_Lab_NPDC), function(x) {
  cat(x, " ", length(unique(Med_Lab_NPDC[[x]])), "\n")
})
```


```{r}
cat("Before duplicates handling: ",dim(Med_Lab_NPDC),"\n")

tmp <- Med_Lab_NPDC %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = npdc_repeat_key,
    values_from = c(npdcd:npdcclass),
    names_sep = "_"
  )
## drop columns if all obs are missing
Med_Lab_NPDC <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(Med_Lab_NPDC),"\n")
```

<br>

### Med_Lab_PDC

```{r}
result <- lapply(names(Med_Lab_PDC), function(x) {
  cat(x, " ", length(unique(Med_Lab_PDC[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(Med_Lab_PDC),"\n")

tmp <- Med_Lab_PDC %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = pdc_repeat_key,
    values_from = c(pdcd:pdcclass),
    names_sep = "_"
  )
## drop columns if all obs are missing
Med_Lab_PDC <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(Med_Lab_PDC),"\n")
```

<br>

### MedHistory_Admit

```{r}
result <- lapply(names(MedHistory_Admit), function(x) {
  cat(x, " ", length(unique(MedHistory_Admit[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(MedHistory_Admit),"\n")

tmp <- MedHistory_Admit %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = AdmitS_Repeat_Key,
    values_from = c(AdmitSRsn),
    names_prefix = "AdmitSRsn_"
  )
## drop columns if all obs are missing
MedHistory_Admit <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(MedHistory_Admit),"\n")
```

<br>

### MedHistory_SurgBF

```{r}
result <- lapply(names(MedHistory_SurgBF), function(x) {
  cat(x, " ", length(unique(MedHistory_SurgBF[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(MedHistory_SurgBF),"\n")

tmp <- MedHistory_SurgBF %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = SurgBF_Repeat_Key,
    values_from = c(SurgBfRsn),
    names_prefix = "SurgBfRsn_"
  )
## drop columns if all obs are missing
MedHistory_SurgBF <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(MedHistory_SurgBF),"\n")
```

<br>

### MedHistory_SurgS

```{r}
result <- lapply(names(MedHistory_SurgS), function(x) {
  cat(x, " ", length(unique(MedHistory_SurgS[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(MedHistory_SurgS),"\n")

tmp <- MedHistory_SurgS %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = SurgS_Repeat_Key,
    values_from = c(SurgSRsn),
    names_prefix = "SurgSRsn_"
  )
## drop columns if all obs are missing
MedHistory_SurgS <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(MedHistory_SurgS),"\n")
```

<br>

### PFMB_SibWAD_or_OtherMemIssue

```{r}
result <- lapply(names(PFMB_SibWAD_or_OtherMemIssue), function(x) {
  cat(x, " ", length(unique(PFMB_SibWAD_or_OtherMemIssue[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(PFMB_SibWAD_or_OtherMemIssue),"\n")

tmp <- PFMB_SibWAD_or_OtherMemIssue %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = FH7SiblingWithAD_OtherMemIssue_Repeat_Key,
    values_from = c(FH7BS1),
    names_prefix = "FH7BS1_"
  )
## drop columns if all obs are missing
PFMB_SibWAD_or_OtherMemIssue <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(PFMB_SibWAD_or_OtherMemIssue),"\n")
```

<br>

### Sleep_MentalAct_OthDevice

```{r}
result <- lapply(names(Sleep_MentalAct_OthDevice), function(x) {
  cat(x, " ", length(unique(Sleep_MentalAct_OthDevice[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(Sleep_MentalAct_OthDevice),"\n")

tmp <- Sleep_MentalAct_OthDevice %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = oth_device_repeat_key,
    values_from = c(SA_OthDevFrq),
    names_prefix = "SA_OthDevFrq_"
  )
## drop columns if all obs are missing
Sleep_MentalAct_OthDevice <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(Sleep_MentalAct_OthDevice),"\n")
```

<br>

### WomenQuestions_BCP

```{r}
result <- lapply(names(WomenQuestions_BCP), function(x) {
  cat(x, " ", length(unique(WomenQuestions_BCP[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(WomenQuestions_BCP),"\n")

tmp <- WomenQuestions_BCP %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = BCP_Repeat_Key,
    values_from = c(BCPName:BCPSide),
    names_sep = "_"
  )
## drop columns if all obs are missing
WomenQuestions_BCP <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(WomenQuestions_BCP),"\n")
```


<br>

### WomenQuestions_Current_HormoneTherapy

```{r}
result <- lapply(names(WomenQuestions_Current_HormoneTherapy), function(x) {
  cat(x, " ", length(unique(WomenQuestions_Current_HormoneTherapy[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(WomenQuestions_Current_HormoneTherapy),"\n")

tmp <- WomenQuestions_Current_HormoneTherapy %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = Cur_HT_Repeat_Key,
    values_from = c(CurMed:CurSide),
    names_sep = "_"
  )
## drop columns if all obs are missing
WomenQuestions_Current_HormoneTherapy <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(WomenQuestions_Current_HormoneTherapy),"\n")
```

<br>

### WomenQuestions_NonPresc_HormoneTherapy

```{r}
result <- lapply(names(WomenQuestions_NonPresc_HormoneTherapy), function(x) {
  cat(x, " ", length(unique(WomenQuestions_NonPresc_HormoneTherapy[[x]])), "\n")
})
```


```{r}
cat("Before duplicates handling: ",dim(WomenQuestions_NonPresc_HormoneTherapy),"\n")

tmp <- WomenQuestions_NonPresc_HormoneTherapy %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = NonPresc_HT_Repeat_Key,
    values_from = c(Horm:HormSide),
    names_sep = "_"
  )
## drop columns if all obs are missing
WomenQuestions_NonPresc_HormoneTherapy <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(WomenQuestions_NonPresc_HormoneTherapy),"\n")
```

<br>

### WomenQuestions_Past_HormoneTherapy

```{r}
result <- lapply(names(WomenQuestions_Past_HormoneTherapy), function(x) {
  cat(x, " ", length(unique(WomenQuestions_Past_HormoneTherapy[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(WomenQuestions_Past_HormoneTherapy),"\n")

tmp <- WomenQuestions_Past_HormoneTherapy %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo),
    names_from = Past_HT_Repeat_Key,
    values_from = c(PastMed:PastStop),
    names_sep = "_"
  )
## drop columns if all obs are missing
WomenQuestions_Past_HormoneTherapy <- tmp[, colSums(!is.na(tmp)) > 0]

cat("After duplicates handling: ",dim(WomenQuestions_Past_HormoneTherapy),"\n")
```


<br>

### WomenQuestions_SERM

```{r}
result <- lapply(names(WomenQuestions_SERM), function(x) {
  cat(x, " ", length(unique(WomenQuestions_SERM[[x]])), "\n")
})
```

```{r}
cat("Before duplicates handling: ",dim(WomenQuestions_SERM),"\n")

tmp <- WomenQuestions_SERM %>%
  pivot_wider(
    id_cols = c(wrapnum, WRAPNo, VisNo,SERMType),
    names_from = SERM_Repeat_Key,
    values_from = c(SERM:SERMSide,-SERMType),
    names_sep = "_"
  )
## drop columns if all obs are missing
WomenQuestions_SERM <- tmp[, colSums(!is.na(tmp)) > 0 | names(tmp) %in% c("SERMType")]

cat("After duplicates handling: ",dim(WomenQuestions_SERM),"\n")
```

<br>
<br>


# Common Columns Detection
```{r}
commonColsDetect(df_names)
```

<br>

## Remove Explanation Column from all Datasets

```{r}
## This column is recording the WRAP webseite link. We remove it to avoid duplicated cases in the merge step
for (df_name in df_names) {
  if ("Explanation" %in% names(get(df_name))) {
    assign(df_name, get(df_name)[ , !(names(get(df_name)) %in% "Explanation") ])
  }
}
```


## Med_Lab_PDC | Med_Lab_NPDC
```{r}
## the common columns between these two data framed are caused by the piovot_wider, mg is the Dose measurement
intersect(names(Med_Lab_PDC),names(Med_Lab_NPDC))
```

```{r}
## I decided to add prefix in each file
colnames(Med_Lab_PDC)[grepl("^mg_", colnames(Med_Lab_PDC))] <- paste0("PDC", colnames(Med_Lab_PDC)[grepl("^mg_", colnames(Med_Lab_PDC))])

colnames(Med_Lab_NPDC)[grepl("^mg_", colnames(Med_Lab_NPDC))] <- paste0("NPDC", colnames(Med_Lab_NPDC)[grepl("^mg_", colnames(Med_Lab_NPDC))])

```

<br>

## IQCode | IADL | CDR | QuickDementia

```{r}
intersect(names(IQCode),names(IADL))
```

```{r}
intersect(names(CDR),names(IADL))
```

```{r}
intersect(names(CDR),names(QuickDementia))
```

```{r}
## add prefix to IQCode and IADL
tNames <- intersect(names(IQCode),names(IADL))[4:8]
colnames(IQCode)[colnames(IQCode) %in% tNames] <-  paste0("IQCode_", colnames(IQCode)[colnames(IQCode) %in% tNames])
colnames(IADL)[colnames(IADL) %in% tNames] <-  paste0("IADL_", colnames(IADL)[colnames(IADL) %in% tNames])

## add prefix to CDR and QuickDementia as well
tNames2 <- tNames[2:5]
colnames(QuickDementia)[colnames(QuickDementia) %in% tNames2] <-  paste0("QuickDementia_", colnames(QuickDementia)[colnames(QuickDementia) %in% tNames2])
colnames(CDR)[colnames(CDR) %in% tNames2] <-  paste0("CDR_", colnames(CDR)[colnames(CDR) %in% tNames2])

## rerun the Common Columns Detection code
commonColsDetect(df_names)
```


<br>

## NeuropsychScores | CompositeScoreCalcValues

```{r}
intersect(names(NeuropsychScores),names(CompositeScoreCalcValues))
```

```{r}
## add prefix to CompositeScoreCalcValues
tNames <- c("flucfl","drraw","stroocw","stroocwe","trlb")
colnames(CompositeScoreCalcValues)[colnames(CompositeScoreCalcValues) %in% tNames] <-  paste0("CompositeScoreCalcValues_", colnames(CompositeScoreCalcValues)[colnames(CompositeScoreCalcValues) %in% tNames])
```

<br>

## IADL | GenHealthDailyAct

```{r}
intersect(names(IADL),names(GenHealthDailyAct))
```

```{r}
## add prefix to GenHealthDailyAct
colnames(GenHealthDailyAct)[colnames(GenHealthDailyAct) == "Money"] <- "GenHealthDailyAct_Money"
```
<br>

## Morbidity_Mortality | GenHealthDailyAct

```{r}
intersect(names(Morbidity_Mortality),names(GenHealthDailyAct))
```

```{r}
colnames(GenHealthDailyAct)[colnames(GenHealthDailyAct) == "MemProb"] <- "GenHealthDailyAct_MemProb"
```

<br>

## NeuropsychScores | Demographics

```{r}
intersect(names(NeuropsychScores),names(Demographics))
```

```{r}
colnames(NeuropsychScores)[colnames(NeuropsychScores) == "interv"] <- "NeuropsychScores_interv"
```

<br>

## PersFamMedBack | Demographics

```{r}
intersect(names(PersFamMedBack),names(Demographics))

## I did not modify anything here since gender has same coding on these two datasets
```

<br>

## Imaging_metadata | fqryStatisticalData

```{r}
intersect(names(Imaging_metadata),names(fqryStatisticalData))
```

```{r}
colnames(Imaging_metadata)[colnames(Imaging_metadata) == "Days_Since_Baseline"] <- "Imaging_metadata_Since_Baseline"
```


<br>
<br>


# Merge Sub-Datasets

Exclude the MIND_Diet_ScoreValues as this exactly the same as the MIND_Diet data frame.

```{r}
# Start with the first dataset

remove <- c("MIND_Diet_ScoredValues") ## this dataset is exactly same as MIND_Diet
df_names2 <- setdiff(df_names,remove)

merged_data <- get(df_names2[1])
cat("Info for", df_names2[1], ":\n")
info(get(df_names2[1]), "wrapnum")

for (i in 2:length(df_names2)) {
  merged_data <- merge_with_info(merged_data, df_names2[i], id_col = "wrapnum")
}
```

```{r}
## clean space: remove everything except final merged data frame
rm(list = setdiff(ls(), "merged_data"))
```

